Training a Gaming Agent on Brainwaves

### 2020-10-19 review2/response.letter.tex

\section*{\fbox{Reviewer \#2 Transcript:}}

Comments to the Author
The authors have satisfactorily addressed my concerns. I have a minor comment regarding Figure 3. It is unclear what the brown bar labelled GrandAvg represents since there is one for each participant. Please clarify. I suggest having another set of bars (a ninth) for the grand average (average over participants) for each of the algorithms used.

\subsection*{\ovalbox{AE General Comments}}
AE's Comments to Author:

Associate Editor
Comments to the Author:
The paper is improved

\begin{quotation}
{\color{blue}
We thanks the AE and all the Reviewers for all the information provided in the comments and for thoughtfully reviewing our manuscript.
}
\end{quotation}

but would benefit from a further update to

@@

Line 50 “The best overall performance is obtained using Logistic Regression.” // also need to discuss the significance of the overall levels of accuracy in Discussion. Is a best performance of 0.672 useful/acceptable? How does it compare to other researchers (if a comparison is possible).

\begin{quotation}
{\color{blue}
On Section IV, third paragraph we tackled this issue.  The overall performance of 0.672 is low.  Although a straightforward comparison could be misleading in the context of the ErrP experiment, we added a brief recount of the values obtained for other similar works.  We also emphasized this point in future works at the end of the Conclusions section.
}
\end{quotation}

Line 42 col 2 “These results are also consistent with their classification ROC curves, shown \sout{on} $<$in$>$ Figures 6 obtained for both subjects, where the area under the curve are close to chance level.” // Why didn’t you show one good ROC curve and one bad one. This would provide the basis for comparison.

\begin{quotation}
{\color{blue}
Based on a previous Reviewer's requests, we added all the ROC curves on the new Figure 6.  We think it is better to allow comparisons between Figures 5,6 and 8.  We are now emphasizing on this figure's caption which ROC curves were showing an effective identification of the ErrP potential and which did not.  We added this in the text, highlighting the important finding that we could not provide an enhancement of the agent performance based on the rewards from subjects where the obtained ROC curves were not good (Section III, page 5).
}
\end{quotation}

@@

Page 5
Figure 5 “Y axis shows the averaged number of steps, while x axis show the number of
experiences used to cumulative train the Q-Table.”
// Why does Fig5 E only have 2 sets of data? Some discussion of the (smaller) change from 1 – 2 would be appropriate (e.g. I Discussion)
Line 51 “No\sout{t} performance gain is evidenced, the agents learn nothing which implies that the reward information is useless.” // useless is not a good choice of wording –provides no value

\begin{quotation}
{\color{blue}
We fixed the inappropriate wording, and appreciate the Reviewer for pointing this out.

Figure 5, for Subject 5 (in the first version of the manuscript it was labeled E) has only two sets of testing data because that Participant (OHC) couldn't complete more sessions.  Hence we used two training sessions to calculate the Q-Table and subsequently performed the run session (i.e. the agent performs movements randomly based on the trained Q-Table for 200 iterations).
On the other hand, we hypothesize that the smaller change obtained from 1-2 for Subject 5 (as well as for Subject 6), may be due to the fact that the ROC curves for those Subjects also display very low classification accuracy.  This could be for reasons related to BCI-illeteracy issues, or the Participant (OHC) not being  concentrated at all during the experiment.  This information was explained in Section III, page 5.
}
\end{quotation}

### 2020-10-19 ErrP.tex
Last version ?

Previous research has explored the usage of RL with reward signals based on brain activity, recorded by an EEG-based BCI system during task execution. The papers \cite{ROBOT-CONTROL-PAPER,Kim2017,Omedes2013} have successfully demonstrated that a robot can be controlled with brain signals from a person who is observing a robot solve a task.  Moreover, a growing number of studies have demonstrated the feasibility of using ErrPs as rewards for RL schemes such as to enhance robotic behaviour~\cite{Luo2019}, to assess air traffic controller's decisions~\cite{Goh2019} or to categorize actions as errors~\cite{Wirth2020}. Other approaches have used these signals as important feedback for human-robot interaction or to implement shared-control strategies~\cite{Schiatti2018,Iturrate2013}.  Additionally, ErrPs have also been used in games as an additional feedback channel that can be explored to improve gaming experience~\cite{Plass-OudeBos2010,kober2018bci}.

@@

In this way, the raw data of a brainwave session is processed into an array of matches where each element is an array of epochs tagged with a number specifying the prediction of the classifier, i.e. if the epoch corresponds to an action that made the agent move further from the goal (hit) or an action that made the agent moves closer to the goal (no-hit). The ErrP is expected to be found in hits. To get the data ready for classification, the stimulus channel is removed to classify the signals using only the EEG data. Each epoch is regularized using a MinMaxScaler, i.e. subtracting the minimum value in the epoch and dividing by the signal peak-to-peak amplitude~\cite{Zhou2019}.  The eight channels are concatenated using the MNE Vectorizer function, which transforms the data matrix into a single array sample. Lastly, this data are used by the classification module as information to train and test a classifier. Five different classification algorithms are used:  Logistic Regression, Multilayer Perceptron with a hidden layer of 100 neurons (i.e. default values for the Scikit-Learn MLPClassifier), Random Forest, KNeighbours with k=3 and finally a linear kernel Support Vector Classifier (i.e. SVM)~\cite{Lotte2018}.

@@

In the case for OHC 5 and 6, the reward information obtained from the brainwaves is not enough to train the agent effectively. Figures \ref{fig:avg_steps} for OHC 5 and 6 show no apparent learning, as the amount of steps to reach the goal doesn't decrease when trained.  These results are also consistent with their classification ROC curves, shown in Figures~\ref{fig:rocsubjects} obtained for both OHCs, where the area under the curve are close to chance level.  Both OHCs have less recorded data from the sessions in comparison to the rest of the OHCs.  This variation in performance for different OHCs has been studied extensively in BCI~\cite{Chavarriaga2014}.  Besides low data samples, there are other reasons affecting the classification accuracy:  cognitive reasons (i.e. the OHC not paying extensively attention to the game dynamics), very low SNR of the ErrP component or even the BCI-illiteracy phenomena where the specific OHC's signals do not contain the expected component response~\cite{Yousefi2019}.

@@

Electroencephalographic signals have high inter-subject variability~\cite{Chavarriaga2014}.  This is evidenced in Figure~\ref{fig:transferlearning} where the agent training is performed with rewards obtained by classifying epochs from one Tester OHC and a classifier that was trained using the brainwaves from a different Trainer OHC.  The figure shows the cumulative variation for all run sessions on the average number of steps required to reach the goal after training the agent with all the available matches from the brainwave session.  Enhancements are shown as negative values.  Only the diagonal of the heatmap matrix shows a clear improvement in terms of the reduction of the required number of steps to reach the goal (averaged per 200 runs) which corresponds to the same information for each OHC shown in Figure~\ref{fig:avg_steps}.  For this transfer learning experiment~\cite{Wu2016}, no performance gain is evidenced, the agents learn nothing and this implies that the reward information provides no value.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. As described in the literature~\cite{Chavarriaga2014}, ErrPs can be used to transmit subjective feedback to a computer.  The collected data show that ErrP signals can in fact be classified and used to train an agent effectively.  This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the human critic. Rewards are generated using the signal processing and classification pipeline, and the Q-Table updates, enhancing the performance of the gaming agent.

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, despite the subjectivity of the ErrP response in terms of the proposed experiment, levels of 90\%  accuracy are reported~\cite{Chavarriaga2014}.  Despite the low accuracy  in ErrP identification, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally, one important aspect of the classification results is the low percentage of false positives (Figure~\ref{fig:confusionmatrix}), showing a high specificity. On the other hand, the percentage of false negatives is generally higher.  Even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Although they may be scarce, accurate rewards are very useful for the RL algorithm.

@@

At the same time, effective agent training depends on the OHC's training data. Results confirm the futility or complexity of using Transfer Learning~\cite{Wu2016}: training a classifier with data obtained from one OHC, and using the same classifier to identify ErrPs for another OHC does not increases the performance of the agent.  Despite that, the rewards generated from different subject's classifiers can be used to train the same Q-Table to improve its performance, which may lead to strategies where the overall performance is enhanced based on the information from different human critics at the same time.  There seems to be an agreement in terms of the subjective interpretation of what may be an appropriate movement to reach the goal.

>>>>

Previous research has explored the usage of RL with reward signals based on brain activity, recorded by an EEG-based BCI system during task execution. The papers \cite{ROBOT-CONTROL-PAPER,Kim2017,Omedes2013} have successfully demonstrated that a robot can be controlled with brain signals from a person who is observing a robot to solve a task.  Moreover, a growing number of studies have demonstrated the feasibility of using ErrPs as rewards for RL schemes such as to enhance robotic behaviour~\cite{Luo2019}, to assess air traffic controller's decisions~\cite{Goh2019} or to categorize actions as errors~\cite{Wirth2020}. Other approaches have used these signals as important feedback for human-robot interaction or to implement shared-control strategies~\cite{Schiatti2018,Iturrate2013}.  Additionally, ErrPs have also been used in games as an additional feedback channel that can be explored to improve gaming experience~\cite{Plass-OudeBos2010,kober2018bci}.

@@

In this way, the raw data of a brainwave session are processed into an array of matches where each element is an array of epochs tagged with a number specifying the prediction of the classifier, i.e. if the epoch corresponds to an action that made the agent move further from the goal (hit) or an action that made the agent moves closer to the goal (no-hit). The ErrP is expected to be found in hits. To get the data ready for classification, the stimulus channel is removed to classify the signals using only the EEG data. Each epoch is regularized using a MinMaxScaler, i.e. subtracting the minimum value in the epoch and dividing by the signal peak-to-peak amplitude~\cite{Zhou2019}.  The eight channels are concatenated using the MNE Vectorizer function, which transforms the data matrix into a single array sample. Lastly, these data are used by the classification module as information to train and test a classifier. Five different classification algorithms are used:  Logistic Regression, Multilayer Perceptron with a hidden layer of 100 neurons (i.e. default values for the Scikit-Learn MLPClassifier), Random Forest, KNeighbours with k=3 and finally a linear kernel Support Vector Classifier (i.e. SVM)~\cite{Lotte2018}.

@@

In the case for OHC 5 and 6, the reward information obtained from the brainwaves is not enough to train the agent effectively. Figures \ref{fig:avg_steps} for OHC 5 and 6 show no apparent learning, as the amount of steps to reach the goal doesn't decrease when trained.  These results are also consistent with the classification ROC curves, shown in Figures~\ref{fig:rocsubjects}, where the area under the curve for OHCs 5 and 6 are close to chance level.  Both OHCs have less recorded data from the sessions in comparison to the rest of the OHCs.  This variation in performance for different OHCs has been studied extensively in BCI~\cite{Chavarriaga2014}.  Besides low data samples, there are other reasons affecting the classification accuracy:  cognitive reasons (i.e. the OHC not paying extensively attention to the game dynamics), very low SNR of the ErrP component or even the BCI-illiteracy phenomena where the specific OHC's signals do not contain the expected component response~\cite{Yousefi2019}.

@@

Electroencephalographic signals have high inter-subject variability~\cite{Chavarriaga2014}.  This is evidenced in Figure~\ref{fig:transferlearning} where the agent training is performed with rewards obtained by classifying epochs from one Tester OHC with a classifier that was trained using the brainwaves from a different Trainer OHC.  The figure shows the cumulative variation for all run sessions on the average number of steps required to reach the goal after training the agent with all the available matches from the brainwave session.  Enhancements are shown as negative values.  Only the diagonal of the heatmap matrix shows a clear improvement in terms of the reduction of the required number of steps to reach the goal (averaged per 200 runs) which corresponds to the same information for each OHC shown in Figure~\ref{fig:avg_steps}.  For this transfer learning experiment~\cite{Wu2016}, no performance gain is evidenced, the agents learn nothing and this implies that the reward information provides no value.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. As described in the literature~\cite{Chavarriaga2014}, ErrPs can be used to transmit subjective feedback to a computer.  The collected data show that ErrP signals can in fact be classified and used to train an agent effectively.  This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the human critic.

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, despite the variability of the ErrP response in terms of different experiments, levels of 90\%  accuracy in identifying it are reported~\cite{Chavarriaga2014}.  Despite the low accuracy  in ErrP identification, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally,  classification results show a low percentage of false positives (Figure~\ref{fig:confusionmatrix}), entailing a high specificity. On the other hand, the percentage of false negatives is generally higher.  Even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Although they may be scarce, accurate rewards are very useful for the RL algorithm.

@@

At the same time, effective agent training depends on the OHC's training data. Results confirm the futility or complexity of using Transfer Learning~\cite{Wu2016}: training a classifier with data obtained from one OHC, and using the same classifier to identify ErrPs for another OHC does not increases the performance of the agent.  Despite that, the rewards generated from different subject's classifiers can be used to train the same Q-Table to improve its performance, which may lead to strategies where the overall performance is enhanced based on the information from different human critics at the same time.  We hypothesize that as long as there are accurate rewards obtained from high specificity classifiers, there is extra information that can be used by the agent to get an improved Q-Table.  Additionally, it seems to be an agreement in terms of the subjective interpretation of what may be an appropriate movement to reach the goal, and different OHCs produce rewards for the same type of movements.
